---
title: "2. patch analysis"
author: "Johannes De Groeve"
date: "1/27/2021"
output: html_document
---

### packages
```{r message=FALSE, error=FALSE, warning=FALSE}
# library(reshape)
# library(ggplot2)
# library(dplyr)
# library(yarrr)
# source('./fncs/21_patch_statistics.R')
```

### data

Dataset including the patch size for every location of roe and red deer location dataset.

* **aniyr**: unique identifier for animal-year
* **study_areas_id**: unique identifier study area as defined in the eurodeer database 
* **sex**: biological sex of the animal
* **animals_id**: unique identifier of animal as defined in the eurodeer database 
* **acquisition_time**: UTC date and time corresponding to a location 
* **x**: x-coordinate of the location using Spatial Reference Identifier SRID 3035 
* **y**: y-coordinate of the location using Spatial Reference Identifier SRID 3035 
* **forest_density**: forest density percentage of the location as defined by the Tree Cover Density Dataset (TCD)
* **corine_land_cover_2012_vector_code**: land cover classification code of the location as defined by Corine Land Cover - Vector (CLC). 
* **timex**: time in utc
* **datex**: date
* **yearx**: year
* **mid_or_noon**: midnight or noon 
* **doyx**: day of the year 
* **min**: minimum doy in dataset
* **max**: maximum doy in dataset 
* **diff_doy**: total number of days as defined by the interval (minimum doy to maximum doy)
* **timex**: time in utc
* **prop**: overall proportion of locations in the forest as defined by TCD 
* **prop_night**: proportion of locations in the forest during the night as defined by TCD 
* **prop_day**: proportion of locations in the forest during the day as defined by TCD 
* **count**: overall number of locations on which calculations are based 
* **count_night**: number of locations during the night on which calculations are based 
* **count_day**: number of locations during the day on which calculations are based 
* **TCD_forest_patch_size**: patch size in squared meter of forest patch of a location as defined by TCD  
* **CLC_forest_patch_size**: patch size in squared meter of forest patch of a location as defined by CLC  
* **TCD_open_patch_size**: patch size in squared meter of open patch of a location as defined by TCD 
* **CLC_open_patch_size**: patch size in squared meter of open patch of a location as defined by CLC 

```{r message=FALSE, error=FALSE}
# ce <- read.csv2('../data/2_patchsize/red.csv',dec = ',')
# ca <- read.csv2('../data/2_patchsize/roe.csv',dec = ',')
# head(ce)
# head(ca)
```

### Apply patch_stat function
```{r message=FALSE, error=FALSE}
# # CALCULATE NUMBER OF PATCHES AND NUMBER OF FIXES IN A PATCH SMALLER THAN 25 HECTARE 
# patch_ce_l <- patch_statistics(ce, patchsize = 250000)
# patch_ca_l <- patch_statistics(ca, patchsize = 250000)
```

### nr of small patches  

First, we calculated the number of patches per individual. 
Then we sum up all the patches over all populations per species. 

```{r message=FALSE, error=FALSE}
# # NR OF PATCHES 
# patch_nr_ce <- patch_ce_l[[1]] 
# patch_nr_ce$species <- 'red'
# patch_nr_ca <- patch_ca_l[[1]] 
# patch_nr_ca$species <- 'roe'
# 
# patch_nr <- rbind(patch_nr_ce,patch_nr_ca)
# patch_nr <- melt(patch_nr, id=c("species","animals_id")) 
# 
# # STATISTICS ABOUT NR OF PATCHES PER INDIVIDUAL OVER ALL POPULATIONS PER SPECIES 
# (r <- group_by(patch_nr, species, variable) %>% 
#     summarise(sum = sum(value), 
#               mean= mean(value), 
#               median=median(value), 
#               max=max(value),
#               min=min(value)))
# barplot(r$sum, 
#         cex.names=0.6,
#         names.arg=c(paste0(r$species, ' ' , r$variable)), 
#         main='total number of patches over all population')
```

### nr of fixes in patches < 25 hectares 

First, we calculated the proportion of fixes that is in a patch smaller than 25 hectare, for each individual. 
Then, we calculate the mean of that proportion over all populations (and thus all individuals)

```{r message=FALSE, error=FALSE}
# patch_fx_ce <- patch_ce_l[[2]]
# patch_fx_ce$species <- 'red'
# patch_fx_ca <- patch_ca_l[[2]] 
# patch_fx_ca$species <- 'roe'
# 
# patch_fx <- rbind(patch_fx_ce,patch_fx_ca)
# (r <- group_by(patch_fx, species, variable) %>% 
#     summarise(mean= mean(Freq), 
#               median=median(Freq), 
#               max=max(Freq),
#               min=min(Freq)))
# barplot(r$mean, 
#         cex.names=0.6,
#         names.arg=c(paste0(r$species, ' ', r$variable)), 
#         main='mean proportion of fixes in small patches (<25ha) per individual over all populations')
# #boxplot(Freq ~ species + variable, data=patch_fx, cex.axis=0.6, main='proportion of fixes in small patches (<25 ha) per individual')
# pirateplot(Freq ~ variable  + species, data=patch_fx, cex.axis=0.6, main='proportion of fixes in small patches (<25 ha) per individual')
```
